### 2024-11-26 paper/paper.qmd
some editting on paper

---
title: "Exploring Urban Patterns in Bicycle Theft: Identifying Key Predictors and Trends"
subtitle: "An Analysis of Bicycle Theft Patterns and Predictors in Toronto"
author: 
  - Tommy Fu
thanks: "Code and data are available at: https://github.com/YichengFu/bike_thefts_analysis.git."
date: today
date-format: long
toc: true
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 paper/paper.qmd
simulation of data

@@
 ---
-title: "Exploring Urban Patterns in Bicycle Theft: Identifying Key Predictors and Trends"
+title: "Understanding Urban Bicycle Theft: Identifying Key Predictors for Risk Assessment"
 subtitle: "An Analysis of Bicycle Theft Patterns and Predictors in Toronto"

### 2024-11-26 scripts/00-simulate_data.R

@@ -1,52 +1,63 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
+# Purpose: Simulates a dataset of bike thefts, including all the variables of interests.
+# Author: Tommy Fu
 # Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
-
+set.seed(432)
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+# Neighborhood names (sample from Toronto or customize further)
+neighborhoods <- c(
+  "Downtown Toronto",
+  "Scarborough",
+  "North York",
+  "Etobicoke",
+  "East York",
+  "York",
+  "West End"
 )
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+# Theft types
+theft_types <- c("Locked", "Unlocked", "Partially Secured")
 
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
+# Simulate dataset
+simulated_data <- tibble(
+  incident_id = 1:500,  # Simulate 500 incidents
+  neighborhood = sample(
+    neighborhoods,
+    size = 500,
     replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
+    prob = c(0.3, 0.2, 0.2, 0.1, 0.1, 0.05, 0.05) # Distribution across neighborhoods
   ),
-  party = sample(
-    parties,
-    size = 151,
+  theft_type = sample(
+    theft_types,
+    size = 500,
     replace = TRUE,
[…]

### 2024-11-26 scripts/01-test_simulated_data.R
sketches drawn

@@ -1,14 +1,11 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
+# Purpose: Tests the structure and validity of the simulated bike thefts data. 
+# Author: Tommy Fu
 # Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####

### 2024-11-26 scripts/01-test_simulated_data.R
simulation data tested

@@ -11,76 +11,126 @@
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Load simulated data
+simulated_data <- tibble(
+  incident_id = 1:500,  # Simulate 500 incidents
+  neighborhood = sample(
+    c("Downtown Toronto", "Scarborough", "North York", "Etobicoke", 
+      "East York", "York", "West End"),
+    size = 500,
+    replace = TRUE
+  ),
+  theft_type = sample(
+    c("Locked", "Unlocked", "Partially Secured"),
+    size = 500,
+    replace = TRUE
+  ),
+  day_of_week = sample(
+    c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
+      "Saturday", "Sunday"),
+    size = 500,
+    replace = TRUE
+  ),
+  time_of_day = sample(
+    c("Morning", "Afternoon", "Evening", "Night"),
+    size = 500,
+    replace = TRUE
+  ),
+  bike_value = round(runif(500, min = 100, max = 3000), 2),  # Random bike values
+  reported_to_police = sample(
+    c(TRUE, FALSE),
+    size = 500,
+    replace = TRUE
+  )
+)
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
+# Test if the dataset was successfully created
+if (exists("simulated_data")) {
+  message("Test Passed: The dataset was successfully created.")
 } else {
-  stop("Test Failed: The dataset could not be loaded.")
+  stop("Test Failed: The dataset could not be created.")
 }
-
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 500 rows
+if (nrow(simulated_data) == 500) {
+  message("Test Passed: The dataset has 500 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 500 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 7 columns
+if (ncol(simulated_data) == 7) {
+  message("Test Passed: The dataset has 7 columns.")
 } else {
[…]

### 2024-11-27 scripts/00-simulate_data.R
written cleaned data and updated reference while working

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates a dataset of bike thefts, including all the variables of interests.
 # Author: Tommy Fu
-# Date: 26 September 2024
+# Date: 26 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: The `tidyverse` package must be installed
 

### 2024-11-27 scripts/01-test_simulated_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated bike thefts data. 
 # Author: Tommy Fu
-# Date: 26 September 2024
+# Date: 27 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded

### 2024-11-27 scripts/02-clean_data.R

@@ -1,44 +1,36 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the raw bike thefts data and drop observations that are missing importatn variables. 
+# Author: Tommy Fu
+# Date: 27 November 2024
+# Contact: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - All the necessary packages must be installed and loaded
+# - 00-simulate_data.R and 01-test_simulated_data.R must have been run
 
 #### Workspace setup ####
-library(tidyverse)
+# Load necessary libraries
+library(dplyr)
+library(tidyr)
+library(arrow)
+library(lubridate) # For date handling
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Load the raw dataset
+bicycle_thefts <- read_csv("data/01-raw_data/Bicycle_Thefts_Open_Data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+# Clean the dataset
+bicycle_thefts_cleaned <- bicycle_thefts %>%
+  # Remove observations where STATUS is "UNKNOWN" or "RECOVERED"
+  filter(!STATUS %in% c("UNKNOWN", "RECOVERED")) %>%
+  # Drop the BIKE_MODEL column due to a high proportion of missing values
+  select(-BIKE_MODEL) %>%
+  # Drop rows with missing values in BIKE_MAKE and BIKE_COST
+  filter(!is.na(BIKE_MAKE), !is.na(BIKE_COST)) %>%
+  # Convert OCC_DATE and REPORT_DATE to Date format
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
[…]

### 2024-11-27 scripts/03-test_analysis_data.R

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Test the analysis data to check for potential problems with cleaning. 
+# Author: Tommy Fu
+# Date: 27 November 2024
+# Contact: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 02-clean_data.R must have been run
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-27 scripts/02-clean_data.R
test analysis data written

@@ -22,8 +22,8 @@ bicycle_thefts <- read_csv("data/01-raw_data/Bicycle_Thefts_Open_Data.csv")
 bicycle_thefts_cleaned <- bicycle_thefts %>%
   # Remove observations where STATUS is "UNKNOWN" or "RECOVERED"
   filter(!STATUS %in% c("UNKNOWN", "RECOVERED")) %>%
-  # Drop the BIKE_MODEL column due to a high proportion of missing values
-  select(-BIKE_MODEL) %>%
+  # Drop the BIKE_MODEL, BIKE_SPEED and BIKE_COLOUR column due to a high proportion of missing values
+  select(-BIKE_MODEL, -BIKE_SPEED, -BIKE_COLOUR) %>%
   # Drop rows with missing values in BIKE_MAKE and BIKE_COST
   filter(!is.na(BIKE_MAKE), !is.na(BIKE_COST)) %>%
   # Convert OCC_DATE and REPORT_DATE to Date format

### 2024-11-27 scripts/03-test_analysis_data.R

@@ -10,59 +10,45 @@
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow) # For reading Parquet files
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
+data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
-
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
-
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
-})
-
-# Test that there are no missing values in the dataset
+# Test that the dataset has no missing values
 test_that("no missing values in dataset", {
-  expect_true(all(!is.na(analysis_data)))
+  expect_true(all(!is.na(data)))
 })
 
-# Test that 'division' contains unique values (no duplicates)
-test_that("'division' column contains unique values", {
-  expect_equal(length(unique(analysis_data$division)), 151)
+# Test that 'occurence_date' and 'report_date' are of Date type
+test_that("'occurence_date' and 'report_date' are Date type", {
+  expect_s3_class(data$occurence_date, "Date")
+  expect_s3_class(data$report_date, "Date")
 })
 
-# Test that 'state' contains only valid Australian state or territory names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
-                  "Tasmania", "Northern Territory", "Australian Capital Territory")
-test_that("'state' contains valid Australian state names", {
-  expect_true(all(analysis_data$state %in% valid_states))
+# Test that 'BIKE_COST' is numeric
+test_that("'BIKE_COST' and 'BIKE_SPEED' are numeric", {
+  expect_type(data$BIKE_COST, "double")
 })
 
-# Test that there are no empty strings in 'division', 'party', or 'state' columns
[…]

### 2024-11-27 scripts/02-clean_data.R
editted pre-requisites

@@ -4,7 +4,7 @@
 # Date: 27 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-# - All the necessary packages must be installed and loaded
+# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate'.
 # - 00-simulate_data.R and 01-test_simulated_data.R must have been run
 
 #### Workspace setup ####

### 2024-11-27 scripts/03-test_analysis_data.R

@@ -4,7 +4,7 @@
 # Date: 27 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-# - The `tidyverse` package must be installed and loaded
+# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate','tidyverse',testthat'.
 # - 02-clean_data.R must have been run
 
 #### Workspace setup ####

### 2024-11-27 scripts/04-exploratory_data_analysis.R
EDA scripts finished

@@ -1,37 +1,114 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Perform Exploratory Data Analysis (EDA) on bike thefts data to explore the data structure and interesting patterns. 
+# Author: Tommy Fu
+# Date: 27 November 2024
+# Contact: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate','tidyverse',testthat'.
+# - 02-clean_data.R must have been run
+# - 03-test_analysis_data must have been run and passed
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(lubridate)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+#### Data structure and summary ####
+# Overview of the dataset
+glimpse(analysis_data)
 
+# Summary statistics
+summary(analysis_data)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+#### Patterns and Trends ####
+# Theft counts over time
+theft_trends <- analysis_data %>%
+  mutate(year_month = floor_date(occurence_date, "month")) %>%
+  count(year_month) %>%
+  arrange(year_month)
 
+# Plot theft trends over time
[…]

### 2024-11-28 scripts/02-clean_data.R
updated cleaning data by adding a dummy variable

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw bike thefts data and drop observations that are missing importatn variables. 
+# Purpose: Cleans the raw bike thefts data, adds high-risk neighborhood variable, and saves the cleaned data.
 # Author: Tommy Fu
 # Date: 27 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate'.
-# - 00-simulate_data.R and 01-test_simulated_data.R must have been run
+# - All the necessary packages must be installed and loaded: 'dplyr', 'tidyr', 'arrow', 'lubridate'.
+# - 00-simulate_data.R and 01-test_simulated_data.R must have been run.
 
 #### Workspace setup ####
 # Load necessary libraries
@@ -22,7 +22,7 @@ bicycle_thefts <- read_csv("data/01-raw_data/Bicycle_Thefts_Open_Data.csv")
 bicycle_thefts_cleaned <- bicycle_thefts %>%
   # Remove observations where STATUS is "UNKNOWN" or "RECOVERED"
   filter(!STATUS %in% c("UNKNOWN", "RECOVERED")) %>%
-  # Drop the BIKE_MODEL, BIKE_SPEED and BIKE_COLOUR column due to a high proportion of missing values
+  # Drop the BIKE_MODEL, BIKE_SPEED, and BIKE_COLOUR columns due to a high proportion of missing values
   select(-BIKE_MODEL, -BIKE_SPEED, -BIKE_COLOUR) %>%
   # Drop rows with missing values in BIKE_MAKE and BIKE_COST
   filter(!is.na(BIKE_MAKE), !is.na(BIKE_COST)) %>%
@@ -31,6 +31,20 @@ bicycle_thefts_cleaned <- bicycle_thefts %>%
     occurence_date = as_date(OCC_DATE, format = "%m/%d/%Y %I:%M:%S %p"),
     report_date = as_date(REPORT_DATE, format = "%m/%d/%Y %I:%M:%S %p")
   )
+
+#### Identify High-Risk Neighborhoods ####
+# Count thefts by neighborhood
+top_neighborhoods <- bicycle_thefts_cleaned %>%
+  count(NEIGHBOURHOOD_140, sort = TRUE) %>%
[…]

### 2024-11-28 scripts/05-model_data.R
model set up and reference updated

@@ -1,37 +1,39 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Built Models for logistic regression model to predict thefts in high-risk neighborhoods.
+# Author: Tommy Fu
+# Date: 27 November 2024
+# Contact: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate','tidyverse',testthat'.
+# - 02-clean_data.R must have been run
+# - 03-test_analysis_data must have been run and passed
+# - 04-exploratory_data_analysis.R must have been run
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 
+
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# ensure outcome consistency by setting seed
+set.seed(432)
 
+# get the reduced dataset of only 1000 randomly selected data entries
+data_reduced <- 
+  analysis_data |> 
+  slice_sample(n = 1000)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+#### Logistic Regression Model ####
+high_risk_model <- stan_glm(
+  formula = is_high_risk_neighborhood ~ BIKE_COST + PREMISES_TYPE + LOCATION_TYPE + OCC_HOUR,
+  data = data_reduced,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
[…]

### 2024-11-28 paper/paper.qmd
EDA scripts finished
paper intro editted some reference packages changed

--- 2024-11-26 paper/paper.qmd
@@
 ---
-title: "Exploring Urban Patterns in Bicycle Theft: Identifying Key Predictors and Trends"
-subtitle: "An Analysis of Bicycle Theft Patterns and Predictors in Toronto"
+title: "Understanding Bicycle Theft in High-Risk Neighborhoods: Key Predictors for Targeted Risk Assessment"
+subtitle: "An Analysis of Theft Patterns and Risk Factors in Toronto's High-Theft Areas"
 author: 
@@
 toc: true
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study explores patterns of bicycle theft in Toronto, focusing on thefts in high-risk neighborhoods using logistic regression.
+The analysis identifies bike cost, premises type, location type, and time of occurrence as significant predictors of thefts in high-risk neighborhoods.
+Results show that high-cost bikes are disproportionately stolen in public spaces and during evening hours, with specific neighborhoods exhibiting elevated theft rates.
+These findings highlight risk factors and provide actionable insights for targeted interventions, urban planning, and improved security measures in vulnerable areas."
 format: pdf
@@
 library(tidyverse)
-library(palmerpenguins)
+library(knitr)
+library(arrow)
+library(here)
+library(rstanarm)
+library(modelsummary)
+
+raw_data <- read_csv(here::here("data/01-raw_data/Bicycle_Thefts_Open_Data.csv"))
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 ```
@@
 
-Overview paragraph
+## Overview
+Bicycle theft is a persistent urban problem with significant economic and emotional impacts on individuals and communities.
+In cities like Toronto, where cycling plays an increasingly vital role in promoting sustainable transportation, understanding the factors contributing to bicycle theft is essential for designing effective preventive measures.
+While previous studies have examined general theft trends, few have focused specifically on identifying patterns in high-risk neighborhoods, where targeted interventions could have the greatest impact.
+In this paper the bike thefts data from Toronto Police Open Data will be utilized .
+This paper seeks to address this gap by exploring the spatial, temporal, and contextual factors associated with bicycle thefts in Toronto.
 
-Estimand paragraph
+## Estimand 
+The estimand of this study is the likelihood of thefts occurring in high-risk neighborhoods compared to others, given key predictors such as bike cost, premises type, location type, and time of theft.
+The estimand focuses on understanding the characteristics of thefts in these neighborhoods, allowing us to identify significant factors that differentiate high-risk areas from others.
+By using a logistic regression model, the analysis aims to quantify these relationships and provide actionable insights.
 
-Results paragraph
+## Results Summary
+The analysis highlights significant relationships between theft occurrences in high-risk neighborhoods and predictors such as bike cost, premises type, and time of day.
+Findings suggest that high-cost bicycles are frequently targeted in public spaces during evening hours, with some neighborhoods experiencing disproportionately higher theft rates.
 
-Why it matters paragraph
+## Why this paper matters
+This research provides actionable knowledge to enhance theft prevention in urban settings.
+By identifying the circumstances under which thefts are more likely to occur, the study offers evidence to inform targeted interventions, such as improved security measures in vulnerable areas.
+These findings contribute to creating safer urban spaces for cyclists.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+## Paper Structure: 
+The remainder of this paper is structured as follows.
+In @sec-data, the overview of the data used in this study and the variables of interests will be introduced.
+Further the data normalization will be discussed in details.
+@sec-model illustrates the Bayesian logistic regression model built in our analysis, some details include model set up, assumptions and justification.
+@sec-result highlights the result of the model visualizing using tables and graphs.
+Lastly, @sec-discussion contains discussion of the analysis based on findings, the limitations of the model and the suggestion for future research.
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+Our data is sourced from Toronto Police Open Data[@toronto_bicycle_thefts] specifically on the bike thefts dataset which captures details in bike thefts reported in Great Toronto Area.
+The bike thefts dataset including details about the bicycles, such as cost, make, and type, as well as information about the thefts, such as the date, time, location type, premises type, and neighborhood.
+Geospatial data, including latitude and longitude, is also included to allow for spatial analysis.
+A sample of the data variables can be find @tbl-rawdatalist in the appendix.
 
@@
 Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
+## Data Processing
 ## Outcome variables
+## Predictor Variable
 
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#ggplot(penguins, aes(x = island, fill = species)) +
+  #geom_bar(alpha = 0.8) +
+  #scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                 #   guide = "none") +
+  #theme_minimal() +
+  #facet_wrap(~species, ncol = 1) +
+  #coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+#analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+#analysis_data |> 
+  #ggplot(aes(x = width, y = length)) +
+  #geom_point(alpha = 0.8) +
+  #theme_minimal() +
+  #labs(x = "Wing width (mm)",
+   #    y = "Wing length (mm)")
 ```
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-result}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
+## Raw data variables
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-rawdatalist
+#| tbl-cap: "The list of variables and a sample value from the raw data"
+#| warning: false
+#### Create Variable List with Sample Values ####
+variable_list <- tibble(
+  `Variable Name` = names(raw_data),
+  `Sample Value` = sapply(raw_data[1, ], as.character) # Extract the first row as sample values
+)
+
+#### Display the Table in Appendix ####
+variable_list %>%
+  kable(
+    format = "markdown",
+    caption = "List of Variables in the Raw Dataset with Sample Values"
+  )
+
+```
+
 # Model details {#sec-model-details}
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+#pp_check(first_model) +
+ #theme_classic() +
+ # theme(legend.position = "bottom")
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+#posterior_vs_prior(first_model) +
+#  theme_minimal() +
+#  scale_color_brewer(palette = "Set1") +
+ # theme(legend.position = "bottom") +
+ # coord_flip()
 ```

### 2024-11-28 scripts/02-clean_data.R

@@ -9,8 +9,7 @@
 
 #### Workspace setup ####
 # Load necessary libraries
-library(dplyr)
-library(tidyr)
+library(tidyverse)
 library(arrow)
 library(lubridate) # For date handling
 

### 2024-11-30 paper/paper.qmd
edits on the data section

@@
 ## Measurement
-
+	This study uses data from Toronto’s open data portal, which provides detailed records of bicycle theft incidents across the city.
+The dataset includes key variables related to theft characteristics, location details, and bike attributes, allowing for a comprehensive analysis of theft patterns in high-risk neighborhoods.
+The dataset focuses exclusively on stolen bicycles, making it well-suited for examining the spatial and contextual factors influencing thefts.
+While other datasets, such as police crime reports or neighborhood demographic data, could provide additional context, they were not used due to limited accessibility or lack of detailed bike-specific information.
 Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

### 2024-12-01 paper/paper.qmd
edits made on data section

--- 2024-11-28 paper/paper.qmd
@@
 library(modelsummary)
+library(forcats)
+library(sf)
 
@@
 ## Results Summary
-The analysis highlights significant relationships between theft occurrences in high-risk neighborhoods and predictors such as bike cost, premises type, and time of day.
-Findings suggest that high-cost bicycles are frequently targeted in public spaces during evening hours, with some neighborhoods experiencing disproportionately higher theft rates.
+The findings of this study reveal significant links between theft occurrences in high-risk neighborhoods and factors such as bike cost, premises type, and time of day.
+High-cost bicycles are more frequently stolen, reflecting their appeal as valuable targets.
+Public spaces, including streets and parks, show higher odds of theft compared to more secure environments like houses or garages, likely due to easier access and fewer security measures.
+
+Thefts are more likely to occur during evening hours compared to morning or daytime periods, a trend that aligns with reduced visibility and activity levels during these hours.
+Additionally, some neighborhoods consistently report higher theft rates, pointing to localized factors such as infrastructure, socioeconomic conditions, or enforcement levels.
+These findings suggest the need for tailored strategies, including improved lighting, enhanced surveillance, and increased public awareness, to reduce theft risks in vulnerable areas and times.
 
@@
 
-Our data is sourced from Toronto Police Open Data[@toronto_bicycle_thefts] specifically on the bike thefts dataset which captures details in bike thefts reported in Great Toronto Area.
-The bike thefts dataset including details about the bicycles, such as cost, make, and type, as well as information about the thefts, such as the date, time, location type, premises type, and neighborhood.
-Geospatial data, including latitude and longitude, is also included to allow for spatial analysis.
-A sample of the data variables can be find @tbl-rawdatalist in the appendix.
+Our dataset is sourced from Toronto Police Open Data [@toronto_bicycle_thefts], specifically the "Bicycle Thefts" dataset, which provides comprehensive details on bicycle thefts reported across the Greater Toronto Area.
+This dataset includes variables that capture information about the stolen bicycles, such as their reported cost, make, and type, as well as contextual details about the thefts, including the date, time, location type, premises type, and neighborhood.
+Additionally, geospatial data, including latitude and longitude coordinates, enables spatial analysis to explore patterns in theft occurrences across different areas.
+Heatmap of Toronto area by theft counts will be shown in [@sec-data-visualization] using shapefile data sourced from University of Toronto Map Library [@trt_shape].
+
+The dataset’s level of granularity allows for an in-depth examination of theft dynamics, facilitating an analysis of how temporal, spatial, and contextual factors interact to influence theft risk.
+For example, information on location and premises type provides insights into whether thefts are more common in public or private spaces, while bike cost highlights economic factors associated with thefts.
+A detailed list of the variables, along with sample values, is provided in the appendix (@tbl-rawdatalist) to offer additional context for understanding the dataset.
+This structured data enables a robust analysis aimed at identifying patterns and predictors of bicycle theft across Toronto.
 
 ## Measurement 
+This study utilizes data from Toronto’s open data portal, specifically focusing on detailed records of bicycle theft incidents reported across the city.
+The dataset includes a rich variety of variables that capture theft characteristics, such as the reported value of the bike, the date and time of the theft, and the theft’s status, alongside spatial details like the neighborhood, premises type, and location type.
+Additionally, geospatial coordinates, including latitude and longitude, enable detailed mapping and spatial analysis of theft hotspots.
+These features make the dataset well-suited for understanding theft dynamics in high-risk neighborhoods, where theft patterns may differ based on socioeconomic, environmental, or infrastructural factors.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset focuses on stolen bicycles, eliminating the noise of other types of crimes and allowing for a more precise examination of the contextual and temporal factors influencing bicycle thefts.
+Key variables, such as BIKE_COST, provide insights into economic patterns of theft, while OCC_DATE and OCC_HOUR allow for the identification of temporal trends.
+Variables like LOCATION_TYPE and PREMISES_TYPE add further depth by categorizing thefts based on their environmental and spatial contexts, helping to discern whether certain locations, such as streets, parks, or garages, are more vulnerable than others.
+
+While other datasets, such as police crime reports or neighborhood demographic data, could have supplemented this analysis, their limited accessibility and lack of detailed information on bike-specific incidents made them unsuitable for the current study.
+The chosen dataset’s specificity and granularity ensure that the analysis remains focused and relevant to the objective of understanding theft patterns in high-risk neighborhoods.
+These features enable a rigorous exploration of the relationships between bike attributes, theft characteristics, and spatial factors, supporting the development of targeted interventions to reduce theft risks in vulnerable areas.
+
 ## Data Processing
+The raw dataset, sourced from Toronto’s open data portal, underwent comprehensive processing steps to ensure it was accurate, relevant, and ready for analysis.
+One of the initial steps involved addressing missing values.
+Variables with a high proportion of missing data, such as BIKE_MODEL, BIKE_SPEED, and BIKE_COLOUR, were removed as they offered limited analytical value.
+Additionally, observations with missing critical values like BIKE_COST or BIKE_MAKE were filtered out to maintain data completeness and integrity.
+This ensured the dataset included only records with sufficient detail for analysis.
+
+The dataset was further refined by excluding theft incidents where the STATUS variable was "UNKNOWN" or "RECOVERED," as these cases were not directly relevant to the study’s focus on thefts.
+Temporal variables, including OCC_DATE (the occurrence date of the theft) and REPORT_DATE (the date the theft was reported), were standardized to a uniform Date format, facilitating the analysis of trends over time, including seasonal and hourly patterns.
+
+A crucial variable, is_high_risk_neighborhood, was constructed to flag thefts occurring in the top 10 neighborhoods with the highest theft frequencies.
+This variable was derived by counting incidents per neighborhood and identifying the areas most affected by theft, enabling the study to focus on high-risk locations.
+The geospatial coordinates LONG_WGS84 and LAT_WGS84 were retained to enable mapping and spatial analysis of theft incidents, providing a foundation for visualizing theft hotspots.
+Contextual variables, such as LOCATION_TYPE (e.g., street, park, store) and PREMISES_TYPE (e.g., house, garage, public area), were also preserved to explore environmental factors influencing theft risk.
+
+The cleaned dataset was saved as a Parquet file, chosen for its efficient storage and compatibility with downstream modeling and visualization workflows.
+This format allowed for fast read-write operations and seamless integration with analysis tools.
+The cleaning and preparation steps ensured that the dataset was robust and aligned with the study’s goal of identifying theft patterns and predictors in high-risk neighborhoods.Packages used in this paper are 'tidyverse' [@tidyverse], 'here' [@here],'arrow' [@arrow], 'lubridate'[@lubridate], 'testthat' [@testthat], 'rstanarm' [@rstanarm], 'knitr' [@knitr], and 'forcats' [@forcats].
+
 ## Outcome variables
-## Predictor Variable
+The dataset includes several important variables that serve as predictors in this study.
+These include **`BIKE_COST`**, a numeric variable representing the reported value of stolen bicycles, which is crucial for understanding how bike value influences theft patterns.
+**`PREMISES_TYPE`** and **`LOCATION_TYPE`** are categorical variables providing contextual details about where thefts occurred, such as public spaces or residential areas, and their environmental settings like streets or parks.
+Temporal details are captured through variables such as **`OCC_HOUR`**, representing the hour of the day the theft occurred, and **`OCC_DATE`**, which allows for trends and seasonal patterns to be explored.
+Geographic variables like **`NEIGHBOURHOOD_140`** and the corresponding longitude and latitude coordinates provide spatial context, enabling an examination of how theft patterns vary across Toronto neighborhoods.
+Together, these variables form the foundation for identifying significant factors associated with bicycle theft patterns.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-cleandata
+#| tbl-cap: "Preview of the Cleaned Data"
+#| warning: false
+#### Display the first few rows ####
+analysis_data %>%
+  slice_head(n = 5) %>% # Display the first 5 rows
+  select(BIKE_COST, OCC_DATE, LOCATION_TYPE, PREMISES_TYPE, STATUS) %>%
+  kable(
+    format = "markdown",
+  )
+```
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+## Predictor variables
+The variable `is_high_risk_neighborhood` is a binary indicator designed to classify thefts based on whether they occurred in one of the top 10 neighborhoods with the highest theft frequencies.
+This variable was constructed to focus the analysis on areas most impacted by bicycle theft.
+A value of 1 indicates that the theft occurred in a high-risk neighborhood, while 0 indicates all other neighborhoods.
+This classification highlights localized clusters of theft activity and enables the study to identify key predictors of theft in these vulnerable areas.
+
+### Distribution of High-Risk Neighborhoods
+
+The majority of reported bicycle thefts occur in a small subset of neighborhoods, demonstrating significant clustering in high-risk areas.
+Figure @fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
 #| echo: false
+#| eval: true
+#| label: fig-riskdistribution
+#| warning: false
+# Graph for High-Risk Neighborhood Distribution
+analysis_data %>%
+  mutate(Category = ifelse(is_high_risk_neighborhood == 1, "High-Risk", "Other")) %>%
+  count(Category, PREMISES_TYPE) %>%
+  group_by(PREMISES_TYPE) %>%
+  mutate(Proportion = n / sum(n)) %>%
+  ungroup() %>%
+  mutate(PREMISES_TYPE = fct_reorder(PREMISES_TYPE, Proportion, .fun = max, .desc = TRUE)) %>% # Reorder by max proportion
+  ggplot(aes(x = PREMISES_TYPE, y = Proportion, fill = Category)) +
+  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
+  labs(
+    title = "Proportion of Thefts by Premises Type and Neighborhood Risk",
+    x = "Premises Type",
+    y = "Proportion",
+    fill = "Neighborhood Type"
+  ) +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_manual(values = c("High-Risk" = "red", "Other" = "gray"))
 
-#ggplot(penguins, aes(x = island, fill = species)) +
-  #geom_bar(alpha = 0.8) +
-  #scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                 #   guide = "none") +
-  #theme_minimal() +
-  #facet_wrap(~species, ncol = 1) +
-  #coord_flip()
+```
+The bar chart shows that high-risk neighborhoods account for a significant proportion of thefts, despite representing only 10 neighborhoods out of the total.
+This concentration highlights the importance of focusing on these areas for targeted interventions.
+
+### Characteristics of High-Risk Neighborhoods
+Table @fig-highrisk_summary provides a summary of the top 10 high-risk neighborhoods, including the total number of thefts recorded in each.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-highrisk_summary
+#| warning: false
+# Ranked Bar Plot for Top 10 High-Risk Neighborhoods
+analysis_data %>%
+  filter(is_high_risk_neighborhood == 1) %>%
+  count(NEIGHBOURHOOD_140, sort = TRUE) %>%
+  slice_max(order_by = n, n = 10) %>% # Specify 'order_by' as 'n'
+  ggplot(aes(x = reorder(NEIGHBOURHOOD_140, n), y = n, fill = n)) +
+  geom_bar(stat = "identity", width = 0.7) +
+  coord_flip() +
+  labs(
+    title = "Top 10 High-Risk Neighborhoods by Theft Count",
+    x = "Neighborhood",
+    y = "Number of Thefts"
+  ) +
+  scale_fill_gradient(low = "blue", high = "red") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
+## Data Visualizations {#sec-data-visualization}
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
 #| echo: false
@@
 #| message: false
+#| label: thefts_map
 
-#analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Load Toronto neighborhood shapefile
+neighbourhoods_sf <- st_read(here::here("data/01-raw_data/Neighbourhoods - 4326.shp"),quiet = TRUE)
 
-#analysis_data |> 
-  #ggplot(aes(x = width, y = length)) +
-  #geom_point(alpha = 0.8) +
-  #theme_minimal() +
-  #labs(x = "Wing width (mm)",
-   #    y = "Wing length (mm)")
+# Load and process the bicycle theft data
+theft_data <- read.csv(here::here("data/01-raw_data/Bicycle_Thefts_Open_Data.csv"))
+
+# Count the number of thefts by neighborhood code, excluding "NSA"
+theft_counts <- theft_data %>%
+  filter(HOOD_158 != "NSA") %>%
+  filter(STATUS != "Recovered") %>%
+  group_by(HOOD_158) %>%
+  summarize(theft_count = n())
+
+# Merge theft counts with neighborhood shapefile
+neighbourhoods_sf <- neighbourhoods_sf %>%
+  left_join(theft_counts, by = c("AREA_LO6" = "HOOD_158"))
+
+# Replace NA theft_count with 0 for neighborhoods with no recorded thefts
+neighbourhoods_sf$theft_count[is.na(neighbourhoods_sf$theft_count)] <- 0
+
+# Plot the map, suppressing warnings
+suppressWarnings({
+  ggplot(data = neighbourhoods_sf) +
+    geom_sf(aes(fill = theft_count), color = "#B0B0B0", linewidth = 0.3) +  # Use linewidth instead of size
+    scale_fill_viridis_c(option = "plasma", name = "Theft Count") +  # Use a color scale for counts
+    labs(title = "Bicycle Theft Count by Neighborhood in Toronto",
+         x = "Longitude", y = "Latitude") +
+    theme_minimal() +
+    theme(
+      plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
+      axis.title = element_text(size = 14),
+      legend.title = element_text(size = 12),
+      legend.text = element_text(size = 10)
+    )
+})
 ```
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
 

### 2024-12-01 paper/paper.qmd
edits on data section and reference updated

@@
 This format allowed for fast read-write operations and seamless integration with analysis tools.
-The cleaning and preparation steps ensured that the dataset was robust and aligned with the study’s goal of identifying theft patterns and predictors in high-risk neighborhoods.Packages used in this paper are 'tidyverse' [@tidyverse], 'here' [@here],'arrow' [@arrow], 'lubridate'[@lubridate], 'testthat' [@testthat], 'rstanarm' [@rstanarm], 'knitr' [@knitr], and 'forcats' [@forcats].
+The cleaning and preparation steps ensured that the dataset was robust and aligned with the study’s goal of identifying theft patterns and predictors in high-risk neighborhoods.Packages used in this paper are 'tidyverse' [@tidyverse], 'here' [@here],'arrow' [@arrow], 'lubridate'[@lubridate], 'testthat' [@testthat], 'rstanarm' [@rstanarm],'sf'[@sf], 'knitr' [@knitr], and 'forcats' [@forcats].
 

### 2024-12-02 scripts/05-model_data.R
changed model name into informative name

@@ -36,4 +36,4 @@ high_risk_model <- stan_glm(
 )
 
 #### Save model ####
-saveRDS(high_risk_model,file = "models/first_model.rds")
+saveRDS(high_risk_model,file = "models/bike_theft_risk_model.rds")

### 2024-12-02 paper/paper.qmd
model section editted

@@
 ## Overview
+
 Bicycle theft is a persistent urban problem with significant economic and emotional impacts on individuals and communities.
@@
 ## Estimand
+
 The estimand of this study is the likelihood of thefts occurring in high-risk neighborhoods compared to others, given key predictors such as bike cost, premises type, location type, and time of theft.
@@
 ## Results Summary
+
 The findings of this study reveal significant links between theft occurrences in high-risk neighborhoods and factors such as bike cost, premises type, and time of day.
@@
 ## Why this paper matters
+
 This research provides actionable knowledge to enhance theft prevention in urban settings.
@@
 ## Paper Structure:
+
 The remainder of this paper is structured as follows.
@@
 For example, information on location and premises type provides insights into whether thefts are more common in public or private spaces, while bike cost highlights economic factors associated with thefts.
-A detailed list of the variables, along with sample values, is provided in the appendix (@tbl-rawdatalist) to offer additional context for understanding the dataset.
+A detailed list of the variables, along with sample values, is provided in the appendix (@sec-data-details) (@tbl-rawdatalist) to offer additional context for understanding the dataset.
 This structured data enables a robust analysis aimed at identifying patterns and predictors of bicycle theft across Toronto.
@@
 ## Measurement
+
 This study utilizes data from Toronto’s open data portal, specifically focusing on detailed records of bicycle theft incidents reported across the city.
@@
 ## Data Processing
+
 The raw dataset, sourced from Toronto’s open data portal, underwent comprehensive processing steps to ensure it was accurate, relevant, and ready for analysis.
@@
 ## Outcome variables
+
 The dataset includes several important variables that serve as predictors in this study.
@@
 Together, these variables form the foundation for identifying significant factors associated with bicycle theft patterns.
+
 ```{r}
@@
 ## Predictor variables
+
 The variable `is_high_risk_neighborhood` is a binary indicator designed to classify thefts based on whether they occurred in one of the top 10 neighborhoods with the highest theft frequencies.
@@
 
-### Distribution of High-Risk Neighborhoods
+## Data Visualizations {#sec-data-visualization}
 
 The majority of reported bicycle thefts occur in a small subset of neighborhoods, demonstrating significant clustering in high-risk areas.
-Figure @fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison.
+Figure @fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison.Table @fig-highrisk_summary in the appendix [@sec-data-details]provides a summary of the top 10 high-risk neighborhoods, including the total number of thefts recorded in each.
 
@@
 #| warning: false
-# Graph for High-Risk Neighborhood Distribution
 analysis_data %>%
-  mutate(Category = ifelse(is_high_risk_neighborhood == 1, "High-Risk", "Other")) %>%
-  count(Category, PREMISES_TYPE) %>%
-  group_by(PREMISES_TYPE) %>%
-  mutate(Proportion = n / sum(n)) %>%
-  ungroup() %>%
-  mutate(PREMISES_TYPE = fct_reorder(PREMISES_TYPE, Proportion, .fun = max, .desc = TRUE)) %>% # Reorder by max proportion
-  ggplot(aes(x = PREMISES_TYPE, y = Proportion, fill = Category)) +
-  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
+  filter(is_high_risk_neighborhood == 1) %>%
+  count(NEIGHBOURHOOD_140, PREMISES_TYPE, sort = TRUE) %>%
+  ggplot(aes(x = reorder(NEIGHBOURHOOD_140, n), y = n, fill = PREMISES_TYPE)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  coord_flip() +
   labs(
-    title = "Proportion of Thefts by Premises Type and Neighborhood Risk",
-    x = "Premises Type",
-    y = "Proportion",
-    fill = "Neighborhood Type"
+    title = "Theft Counts by Premises Type in High-Risk Neighborhoods",
+    x = "Neighborhood",
+    y = "Number of Thefts",
+    fill = "Premises Type"
   ) +
-  theme_minimal() +
-  coord_flip() +
-  scale_fill_manual(values = c("High-Risk" = "red", "Other" = "gray"))
+  theme_minimal()
 
 ```
+
 The bar chart shows that high-risk neighborhoods account for a significant proportion of thefts, despite representing only 10 neighborhoods out of the total.
@@
 
-### Characteristics of High-Risk Neighborhoods
-Table @fig-highrisk_summary provides a summary of the top 10 high-risk neighborhoods, including the total number of thefts recorded in each.
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: fig-highrisk_summary
-#| warning: false
-# Ranked Bar Plot for Top 10 High-Risk Neighborhoods
-analysis_data %>%
-  filter(is_high_risk_neighborhood == 1) %>%
-  count(NEIGHBOURHOOD_140, sort = TRUE) %>%
-  slice_max(order_by = n, n = 10) %>% # Specify 'order_by' as 'n'
-  ggplot(aes(x = reorder(NEIGHBOURHOOD_140, n), y = n, fill = n)) +
-  geom_bar(stat = "identity", width = 0.7) +
-  coord_flip() +
-  labs(
-    title = "Top 10 High-Risk Neighborhoods by Theft Count",
-    x = "Neighborhood",
-    y = "Number of Thefts"
-  ) +
-  scale_fill_gradient(low = "blue", high = "red") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-```
-
-## Data Visualizations {#sec-data-visualization}
 ```{r}
@@
 
+This map visualizes bicycle theft counts by neighborhood across Toronto, highlighting spatial patterns in theft frequency.
+The intensity of the color gradient, ranging from dark purple (low theft count) to bright yellow (high theft count), indicates the relative number of bicycle thefts in each neighborhood.
+The visualization reveals that bicycle thefts are highly concentrated in certain neighborhoods, particularly in the downtown core and surrounding areas, as indicated by the bright yellow and orange regions.
+These areas may be more susceptible to theft due to higher population density, greater cycling activity, or specific environmental factors such as the availability of parking infrastructure or security measures.
+This map underscores the spatial disparity in theft incidents and suggests that targeted interventions, such as improved bike parking security and increased awareness campaigns, may be particularly beneficial in high-theft neighborhoods.
+This visualization also provides a foundation for further analysis, such as exploring the relationship between theft hotspots and socioeconomic or infrastructure variables.
+
 # Model {#sec-model}
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Model Overview
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+In this analysis, we utilized a Bayesian logistic regression model to examine the factors associated with bicycle thefts in high-risk neighborhoods.
+The dependent variable is `is_high_risk_neighborhood`, a binary indicator identifying whether a theft occurred in one of the top 10 neighborhoods with the highest theft counts.
+Detailed model diagnostics and background information are available in Appendix [@sec-model-details].
 
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
+The model is specified as follows: 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i \mid \pi_i &\sim \text{Bern}(\pi_i) \\
+\text{logit}(\pi_i) &= \alpha + \beta_1 \times \text{BIKE\_COST}_i 
+    + \beta_2 \times \text{PREMISES\_TYPE}_i \notag \\
+    &\quad + \beta_3 \times \text{LOCATION\_TYPE}_i 
+    + \beta_4 \times \text{OCC\_HOUR}_i \\
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_1, \beta_2, \beta_3, \beta_4 &\sim \text{Normal}(0, 2.5)
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+Here, $y_i$ represents the binary outcome variable indicating whether a theft occurred in a high-risk neighborhood.
+The probability of this event, ($\pi_i$) is modeled using a logistic link function.
+Predictors in the model include the reported BIKE_COST, the PREMISES_TYPE and LOCATION_TYPE of the theft, and the time of day (OCC_HOUR) when the theft occurred.Weakly informative priors were used to regularize the model.
+Specifically, all parameters($\alpha$,$\beta_1$,$\beta_2$,$\beta_3$,$\beta_4$) were assigned normal prior distributions with a mean of 0 and a standard deviation of 2.5.
+Sampling for the model was conducted using Markov Chain Monte Carlo (MCMC) methods implemented in the rstanarm package @rstanarm in R[@citeR].
+To optimize runtime, a random sample of 1000 data entries was used, with a seed of 215 to ensure reproducibility.
+Model diagnostics, including convergence checks and posterior summaries, are presented in Appendix [@sec-model-details].
 
-### Model justification
+## Model justification
+The Bayesian logistic regression was chosen for its suitability for binary outcome variables and its ability to incorporate prior information.
+This approach enables a probabilistic interpretation of results, allowing for uncertainty quantification in parameter estimates.
+The use of weakly informative priors helps stabilize the model and prevents overfitting, particularly given the limited sample size.
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Regarding the predictors, we hypothesize the following relationships:
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+-BIKE_COST: Higher-cost bicycles are more likely to be targeted in thefts due to their resale value and desirability.
+-PREMISES_TYPE and LOCATION_TYPE: Theft risk may be higher in public or semi-public spaces, such as streets and parks, where security measures are limited.
+-OCC_HOUR: Theft likelihood is expected to increase during evening hours when visibility and public activity are reduced.
+
+The Bayesian logistic regression model provides a robust framework for examining these relationships, offering insights into the factors influencing thefts in high-risk neighborhoods.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
-# Additional data details
+# Additional data details {#sec-data-details}
 
 ## Raw data variables
+
 ```{r}
@@
 
+## Data visualization
+@fig-highrisk_summary is the Ranked Bar Plot for Top 10 High-Risk Neighborhoods.
+@fig-thefts_by_day shows the bike thefts count by the time of the day. 
+@fig-thefts_premises_type shows the bike thefts count by different type of premises.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-highrisk_summary
+#| warning: false
+# Ranked Bar Plot for Top 10 High-Risk Neighborhoods
+analysis_data %>%
+  filter(is_high_risk_neighborhood == 1) %>%
+  count(NEIGHBOURHOOD_140, sort = TRUE) %>%
+  slice_max(order_by = n, n = 10) %>% # Specify 'order_by' as 'n'
+  ggplot(aes(x = reorder(NEIGHBOURHOOD_140, n), y = n, fill = n)) +
+  geom_bar(stat = "identity", width = 0.7) +
+  coord_flip() +
+  labs(
+    title = "Top 10 High-Risk Neighborhoods by Theft Count",
+    x = "Neighborhood",
+    y = "Number of Thefts"
+  ) +
+  scale_fill_gradient(low = "blue", high = "red") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-thefts_by_day
+#| warning: false
+# Theft by hour of the day
+analysis_data %>%
+  count(OCC_HOUR) %>%
+  ggplot(aes(x = OCC_HOUR, y = n, fill = n)) +
+  geom_bar(stat = "identity") +
+  scale_fill_gradient(low = "yellow", high = "red") +
+  labs(
+    title = "Theft Counts by Time of Day",
+    x = "Hour of Day",
+    y = "Number of Thefts"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-thefts_premises_type
+#| warning: false
+# Bar plot for premises type
+analysis_data %>%
+  count(PREMISES_TYPE, sort = TRUE) %>%
+  ggplot(aes(x = reorder(PREMISES_TYPE, n), y = n, fill = n)) +
+  geom_bar(stat = "identity") +
+  coord_flip() +
+  scale_fill_gradient(low = "blue", high = "red") +
+  labs(
+    title = "Bicycle Thefts by Premises Type",
+    x = "Premises Type",
+    y = "Number of Thefts"
+  ) +
+  theme_minimal()
+
+```
+
 # Model details {#sec-model-details}
@@
 ## Posterior predictive check
+Figure @fig-post_dist shows the posterior predictive distribution, comparing the observed data (actual theft occurrences in high-risk neighborhoods) with predictions generated by the posterior distribution of our Bayesian logistic regression model.
+The close alignment between observed and predicted data indicates that the model is well-calibrated and effectively captures patterns in the data.
+This provides confidence that the model represents the underlying processes driving theft patterns and is capable of producing reliable inferences.
 
+```{r}
+#| label: fig-post_dist
+#| fig-cap: Posterior distribution for logistic regression model 
+#| echo: false
+#| warning: false
+#| message: false
+
+pp_check() +
+  theme(legend.position = "bottom")
+```
+
+Figure @fig-post_prior compares the posterior and prior distributions of the model parameters.
+For some predictors, such as thefts in high-risk neighborhoods involving specific premises types or during certain times, the posterior distributions shift significantly compared to the priors.
+This indicates that the observed data strongly influenced the parameter estimates, updating our prior beliefs.
+For others, the posterior and prior distributions overlap substantially, suggesting that the observed data aligns well with prior expectations.
 In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

### 2024-12-02 scripts/05-model_data.R
final model edits

@@ -26,7 +26,7 @@ data_reduced <-
   slice_sample(n = 1000)
 
 #### Logistic Regression Model ####
-high_risk_model <- stan_glm(
+bike_theft_risk_model <- stan_glm(
   formula = is_high_risk_neighborhood ~ BIKE_COST + PREMISES_TYPE + LOCATION_TYPE + OCC_HOUR,
   data = data_reduced,
   family = binomial(link = "logit"),
@@ -36,4 +36,4 @@ high_risk_model <- stan_glm(
 )
 
 #### Save model ####
-saveRDS(high_risk_model,file = "models/bike_theft_risk_model.rds")
+saveRDS(bike_theft_risk_model,file = "models/bike_theft_risk_model.rds")

### 2024-12-02 scripts/05-model_data.R
updated final model

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Built Models for logistic regression model to predict thefts in high-risk neighborhoods.
+# Purpose: Build a logistic regression model to predict thefts in high-risk neighborhoods with improved predictor transformations.
 # Author: Tommy Fu
 # Date: 27 November 2024
 # Contact: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-# - All the necessary packages must be installed and loaded 'dplyr','tidyr','arrow','lubridate','tidyverse',testthat'.
+# - All the necessary packages must be installed and loaded: 'dplyr', 'tidyr', 'arrow', 'lubridate', 'tidyverse', 'rstanarm'.
 # - 02-clean_data.R must have been run
 # - 03-test_analysis_data must have been run and passed
 # - 04-exploratory_data_analysis.R must have been run
@@ -12,22 +12,34 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
-
+library(lubridate)
+library(arrow)
 
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-# ensure outcome consistency by setting seed
+# Transform predictors for improved model performance
+analysis_data <- analysis_data %>%
+  # Log-transform BIKE_COST to handle skewness
+  mutate(log_bike_cost = log1p(BIKE_COST)) %>%
+  # Categorize OCC_HOUR into broader time-of-day groups
+  mutate(time_of_day = case_when(
+    OCC_HOUR >= 6 & OCC_HOUR < 12 ~ "Morning",
+    OCC_HOUR >= 12 & OCC_HOUR < 18 ~ "Afternoon",
+    OCC_HOUR >= 18 & OCC_HOUR < 24 ~ "Evening",
+    TRUE ~ "Night"
+  ))
+
+# Ensure consistent outcome by setting a seed
 set.seed(432)
 
-# get the reduced dataset of only 1000 randomly selected data entries
-data_reduced <- 
-  analysis_data |> 
+# Sample a subset of data for faster model fitting
+data_reduced <- analysis_data %>% 
   slice_sample(n = 1000)
 
 #### Logistic Regression Model ####
 bike_theft_risk_model <- stan_glm(
[…]

### 2024-12-02 paper/paper.qmd
Result section edit

@@
 abstract: "This study explores patterns of bicycle theft in Toronto, focusing on thefts in high-risk neighborhoods using logistic regression.
-The analysis identifies bike cost, premises type, location type, and time of occurrence as significant predictors of thefts in high-risk neighborhoods.
+The analysis identifies bike cost, premises type, and time of occurrence as significant predictors of thefts in high-risk neighborhoods.
 Results show that high-cost bikes are disproportionately stolen in public spaces and during evening hours, with specific neighborhoods exhibiting elevated theft rates.
@@
 analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+analysis_data <- analysis_data %>%
+  mutate(log_bike_cost = log1p(BIKE_COST))%>%
+mutate(time_of_day = case_when(
+    OCC_HOUR >= 6 & OCC_HOUR < 12 ~ "Morning",
+    OCC_HOUR >= 12 & OCC_HOUR < 18 ~ "Afternoon",
+    OCC_HOUR >= 18 & OCC_HOUR < 24 ~ "Evening",
+    TRUE ~ "Night"
+  )) %>%
+  # Ensure time_of_day is a factor for modeling
+  mutate(time_of_day = factor(time_of_day, levels = c("Morning", "Afternoon", "Evening", "Night")))
 ```
@@
 
-The estimand of this study is the likelihood of thefts occurring in high-risk neighborhoods compared to others, given key predictors such as bike cost, premises type, location type, and time of theft.
+The estimand of this study is the likelihood of thefts occurring in high-risk neighborhoods compared to others, given key predictors such as bike cost, premises type, and time of theft.
 The estimand focuses on understanding the characteristics of thefts in these neighborhoods, allowing us to identify significant factors that differentiate high-risk areas from others.
@@
 Our dataset is sourced from Toronto Police Open Data [@toronto_bicycle_thefts], specifically the "Bicycle Thefts" dataset, which provides comprehensive details on bicycle thefts reported across the Greater Toronto Area.
-This dataset includes variables that capture information about the stolen bicycles, such as their reported cost, make, and type, as well as contextual details about the thefts, including the date, time, location type, premises type, and neighborhood.
+This dataset includes variables that capture information about the stolen bicycles, such as their reported cost, make, and type, as well as contextual details about the thefts, including the date, time, premises type, and neighborhood.
 Additionally, geospatial data, including latitude and longitude coordinates, enables spatial analysis to explore patterns in theft occurrences across different areas.
-Heatmap of Toronto area by theft counts will be shown in [@sec-data-visualization] using shapefile data sourced from University of Toronto Map Library [@trt_shape].
+Heatmap of Toronto area by theft counts and other informative plots will be shown in [@sec-data-visualization].
+The heatmap utilized shapefile data sourced from University of Toronto Map Library [@trt_shape].
 
 The dataset’s level of granularity allows for an in-depth examination of theft dynamics, facilitating an analysis of how temporal, spatial, and contextual factors interact to influence theft risk.
-For example, information on location and premises type provides insights into whether thefts are more common in public or private spaces, while bike cost highlights economic factors associated with thefts.
+For example, information on premises type provides insights into whether thefts are more common in public or private spaces, while bike cost highlights economic factors associated with thefts.
 A detailed list of the variables, along with sample values, is provided in the appendix (@sec-data-details) (@tbl-rawdatalist) to offer additional context for understanding the dataset.
@@
 
-This study utilizes data from Toronto’s open data portal, specifically focusing on detailed records of bicycle theft incidents reported across the city.
-The dataset includes a rich variety of variables that capture theft characteristics, such as the reported value of the bike, the date and time of the theft, and the theft’s status, alongside spatial details like the neighborhood, premises type, and location type.
+This study utilizes data from Toronto Police open data portal [@toronto_bicycle_thefts], specifically focusing on detailed records of bicycle theft incidents reported across the city.
+The dataset includes a rich variety of variables that capture theft characteristics, such as the reported value of the bike, the date and time of the theft, and the theft’s status, alongside spatial details like the neighborhood and premises type.
 Additionally, geospatial coordinates, including latitude and longitude, enable detailed mapping and spatial analysis of theft hotspots.
@@
 Key variables, such as BIKE_COST, provide insights into economic patterns of theft, while OCC_DATE and OCC_HOUR allow for the identification of temporal trends.
-Variables like LOCATION_TYPE and PREMISES_TYPE add further depth by categorizing thefts based on their environmental and spatial contexts, helping to discern whether certain locations, such as streets, parks, or garages, are more vulnerable than others.
+Variables like PREMISES_TYPE add further depth by categorizing thefts based on their environmental and spatial contexts, helping to discern whether certain locations, such as streets, parks, or garages, are more vulnerable than others.
 
@@
 
-The raw dataset, sourced from Toronto’s open data portal, underwent comprehensive processing steps to ensure it was accurate, relevant, and ready for analysis.
+The raw dataset, sourced from Toronto Police open data portal [@toronto_bicycle_thefts], underwent comprehensive processing steps to ensure it was accurate, relevant, and ready for analysis.
 One of the initial steps involved addressing missing values.
@@
 This ensured the dataset included only records with sufficient detail for analysis.
+More details available in @sec-data-details
 
@@
 The geospatial coordinates LONG_WGS84 and LAT_WGS84 were retained to enable mapping and spatial analysis of theft incidents, providing a foundation for visualizing theft hotspots.
-Contextual variables, such as LOCATION_TYPE (e.g., street, park, store) and PREMISES_TYPE (e.g., house, garage, public area), were also preserved to explore environmental factors influencing theft risk.
+Contextual variables, such as PREMISES_TYPE (e.g., house, garage, public area), were also preserved to explore environmental factors influencing theft risk.
 
@@
 These include **`BIKE_COST`**, a numeric variable representing the reported value of stolen bicycles, which is crucial for understanding how bike value influences theft patterns.
-**`PREMISES_TYPE`** and **`LOCATION_TYPE`** are categorical variables providing contextual details about where thefts occurred, such as public spaces or residential areas, and their environmental settings like streets or parks.
+**`PREMISES_TYPE`** is categorical variables providing contextual details about where thefts occurred, such as public spaces or residential areas, and their environmental settings like streets or parks.
 Temporal details are captured through variables such as **`OCC_HOUR`**, representing the hour of the day the theft occurred, and **`OCC_DATE`**, which allows for trends and seasonal patterns to be explored.
@@
   slice_head(n = 5) %>% # Display the first 5 rows
-  select(BIKE_COST, OCC_DATE, LOCATION_TYPE, PREMISES_TYPE, STATUS) %>%
+  select(BIKE_COST, OCC_DATE, LOCATION_TYPE, PREMISES_TYPE) %>%
   kable(
@@
 The majority of reported bicycle thefts occur in a small subset of neighborhoods, demonstrating significant clustering in high-risk areas.
-Figure @fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison.Table @fig-highrisk_summary in the appendix [@sec-data-details]provides a summary of the top 10 high-risk neighborhoods, including the total number of thefts recorded in each.
+@fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison in top ten risky neighborhoods .[@fig-thefts_map] visualizes bicycle theft counts by neighborhood across Toronto, highlighting spatial patterns in theft frequency.
+More detailed visualizations of the data are located in [@sec-data-details-visualizations]
 
@@
 #| warning: false
+#| fig-cap: "Theft Counts by Premises Type in High-Risk Neighborhoods"
 analysis_data %>%
@@
 #| message: false
-#| label: thefts_map
+#| label: fig-thefts_map
+#| fig-cap: "Bicycle Theft Count by Neighborhood in Toronto"
 
@@
 
-This map visualizes bicycle theft counts by neighborhood across Toronto, highlighting spatial patterns in theft frequency.
 The intensity of the color gradient, ranging from dark purple (low theft count) to bright yellow (high theft count), indicates the relative number of bicycle thefts in each neighborhood.
@@
 y_i \mid \pi_i &\sim \text{Bern}(\pi_i) \\
-\text{logit}(\pi_i) &= \alpha + \beta_1 \times \text{BIKE\_COST}_i 
+\text{logit}(\pi_i) &= \alpha + \beta_1 \times \text{LOG\_BIKE\_COST}_i 
     + \beta_2 \times \text{PREMISES\_TYPE}_i \notag \\
-    &\quad + \beta_3 \times \text{LOCATION\_TYPE}_i 
-    + \beta_4 \times \text{OCC\_HOUR}_i \\
+    &\quad + \beta_3 \times \text{OCC\_HOUR}_i \\
 \alpha &\sim \text{Normal}(0, 2.5) \\
-\beta_1, \beta_2, \beta_3, \beta_4 &\sim \text{Normal}(0, 2.5)
+\beta_1, \beta_2, \beta_3 &\sim \text{Normal}(0, 2.5)
 \end{align}
@@
 The probability of this event, ($\pi_i$) is modeled using a logistic link function.
-Predictors in the model include the reported BIKE_COST, the PREMISES_TYPE and LOCATION_TYPE of the theft, and the time of day (OCC_HOUR) when the theft occurred.Weakly informative priors were used to regularize the model.
-Specifically, all parameters($\alpha$,$\beta_1$,$\beta_2$,$\beta_3$,$\beta_4$) were assigned normal prior distributions with a mean of 0 and a standard deviation of 2.5.
+Predictors in the model include the reported BIKE_COST, the PREMISES_TYPE of the theft, and the time of day (OCC_HOUR) when the theft occurred.Weakly informative priors were used to regularize the model.
+Specifically, all parameters($\alpha$,$\beta_1$,$\beta_2$,$\beta_3$) were assigned normal prior distributions with a mean of 0 and a standard deviation of 2.5.
 Sampling for the model was conducted using Markov Chain Monte Carlo (MCMC) methods implemented in the rstanarm package @rstanarm in R[@citeR].
@@
 -BIKE_COST: Higher-cost bicycles are more likely to be targeted in thefts due to their resale value and desirability.
--PREMISES_TYPE and LOCATION_TYPE: Theft risk may be higher in public or semi-public spaces, such as streets and parks, where security measures are limited.
+-PREMISES_TYPE: Theft risk may be higher in public or semi-public spaces, such as streets and parks, where security measures are limited.
 -OCC_HOUR: Theft likelihood is expected to increase during evening hours when visibility and public activity are reduced.
@@
 #| message: false
+#| include: false
+first_model <-
+  readRDS(file = here::here("models/bike_theft_risk_model.rds"))
+# Ensure consistent outcome by setting a seed
+set.seed(432)
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# Sample a subset of data for faster model fitting
+data_reduced <- analysis_data %>% 
+  slice_sample(n = 1000)
 ```
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Bike Thefts in High Risk Area(n=1000)"
 #| warning: false
 
-modelsummary::modelsummary(
+modelsummary(
   list(
-    "First model" = first_model
+    "Bike Thefts in High Risk Area " = first_model
   ),
-  statistic = "mad",
-  fmt = 2
+  statistic = "mad"
 )
 ```
+
+## Model Validation
 
@@
 
-## Data visualization
+## Data visualization {#sec-data-details-visualizations}
 @fig-highrisk_summary is the Ranked Bar Plot for Top 10 High-Risk Neighborhoods.
@@
 #| warning: false
+#| fig-cap: "Ranked Bar Plot for Top 10 High-Risk Neighborhoods"
 # Ranked Bar Plot for Top 10 High-Risk Neighborhoods
@@
 #| warning: false
+#| fig-cap: "Theft Counts by hour of the day"
 # Theft by hour of the day
@@
 #| warning: false
+#| fig-cap: "Bicycle Thefts by Premises Type"
 # Bar plot for premises type
@@
 
-pp_check() +
+pp_check(first_model) +
   theme(legend.position = "bottom")
@@
 For others, the posterior and prior distributions overlap substantially, suggesting that the observed data aligns well with prior expectations.
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
 
 ```{r}
-#| eval: true
+#| label: fig-post_prior
+#| fig-cap: "Comparison of Posterior and Prior Distributions"
 #| echo: false
+#| warning: false
 #| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-width: 14
+#| fig-height: 12
 
-#pp_check(first_model) +
- #theme_classic() +
- # theme(legend.position = "bottom")
+posterior_vs_prior(first_model, size = 0.2) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_text(size = 12, face = "bold"),
+    legend.text = element_text(size = 10),
+    plot.margin = margin(15, 15, 30, 15), # Increase spacing around plot
+    axis.text.y = element_text(size = 8), # Smaller y-axis labels for long parameter names
+    axis.text.x = element_text(size = 9),
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(size = 12, hjust = 0.5),
+    axis.title = element_text(size = 12, face = "bold") # Emphasize axis titles
+  ) +
+  guides(color = guide_legend(ncol = 2)) + # Adjust legend layout for compact display
+  labs(
+    title = "Comparison of Posterior and Prior Distributions",
+    subtitle = "Shifts in parameter distributions after observing data",
+    x = "Parameter Value",
+    y = "Parameter"
+  ) +
+  coord_flip()
 
-#posterior_vs_prior(first_model) +
-#  theme_minimal() +
-#  scale_color_brewer(palette = "Set1") +
- # theme(legend.position = "bottom") +
- # coord_flip()
 ```
 
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+## Markov chain Monte Carlo Convergence Check
+Figures @fig-trace1 and @fig-trace2 display the trace plots for our Bayesian logistic regression model, providing a detailed assessment of the convergence of the Markov Chain Monte Carlo (MCMC) sampling process.
+These plots visualize the parameter chains across iterations for key predictors selected based on their relevance to theft patterns.
+Specifically, parameters like **Intercept**, **Log Bike Cost**, and representative **Premises Type** are shown due to their strong potential to influence theft likelihood.
+The trace plot reveal that the chains oscillate horizontally, overlap between chains, and lack any noticeable divergence, indicating stable sampling and effective convergence.
+This stability ensures that the model provides reliable estimates for analyzing theft risk factors in high-risk neighborhoods.
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| label: fig-trace1
+#| fig-cap: "Trace plot of premises type"
+#| fig-subcap: ["Trace Plot of Commercial Premise", "Trace Plot of Educational Premise","Trace Plot of House Premise","Trace Plot of Outside Premise","Trace Plot of Transit Premise","Trace Plot of Other Premise"]
 #| layout-ncol: 2
 
-plot(first_model, "trace")
+# Generate trace plot for premises type
+plot(first_model, "trace", "PREMISES_TYPECommercial")
+plot(first_model, "trace", "PREMISES_TYPEEducational")
+plot(first_model, "trace", "PREMISES_TYPEHouse")
+plot(first_model, "trace", "PREMISES_TYPEOutside")
+plot(first_model, "trace", "PREMISES_TYPETransit")
+plot(first_model, "trace", "PREMISES_TYPEOther")
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trace2
+#| fig-cap: "Trace plots for time_of_day, intercept, and log_bike_cost"
+#| fig-subcap: ["Trace Plot for Intercept", "Trace Plot for log Bike Cost", "Trace Plot for Morning","Trace Plot for Evening","Trace Plot for Night"]
+#| layout-ncol: 2
+
+# Parameters to include in the trace plots
+parameters <- c("(Intercept)", "log_bike_cost", "time_of_dayMorning", "time_of_dayEvening", "time_of_dayNight")
+
+# Generate trace plots for each parameter
+plot(first_model, "trace", "(Intercept)")
+plot(first_model, "trace", "log_bike_cost")
+plot(first_model, "trace", "time_of_dayMorning")
+plot(first_model, "trace", "time_of_dayEvening")
+plot(first_model, "trace", "time_of_dayNight")
+
+```
+
+Figure @fig-rhat presents the Rhat diagnostic for the model parameters, a key measure of convergence in MCMC sampling.
+The Rhat statistic compares the variability within each chain to the variability between chains, providing insight into whether the chains have mixed well.
+In this model, all Rhat values are very close to 1 and remain well below the threshold of 1.05, indicating that the chains have converged effectively.
+This result reinforces the reliability of the posterior estimates and confirms that the sampling process has stabilized, ensuring the model’s robustness for inference.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-rhat
+#| fig-cap: "Rhat plot"
 
 plot(first_model, "rhat")
+```
+
+### 90% Credibility Interval
+
+Figure @fig-modelresults1 displays the 90% credible intervals for the predictors in the Bayesian logistic regression model.
+These intervals illustrate the uncertainty surrounding each parameter estimate, allowing us to assess their significance and direction of effect.
+Parameters whose intervals do not cross zero are considered statistically significant, as they demonstrate a consistent association with the likelihood of bike theft in high-risk neighborhoods.
+For example, certain premises types show strong, consistent effects, while others, such as specific time-of-day variables, have wider intervals reflecting greater uncertainty.
+This visualization provides a clear summary of the model's results and highlights the most impactful predictors of theft risk.
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-modelresults1
+#| fig-cap: "Credible intervals for predictors of positive poverty status"
+
+modelplot(first_model, conf_level = 0.90, size = 0.2) +
+  labs(x = "90% credibility interval") +
+  theme(axis.text.y = element_text(size = 6))
 ```

### 2024-12-02 paper/paper.qmd
result section editted

@@
 The majority of reported bicycle thefts occur in a small subset of neighborhoods, demonstrating significant clustering in high-risk areas.
-@fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison in top ten risky neighborhoods .[@fig-thefts_map] visualizes bicycle theft counts by neighborhood across Toronto, highlighting spatial patterns in theft frequency.
+@fig-riskdistribution illustrates the distribution of thefts by risk classification with different premises types comparison in top ten risky neighborhoods .
+[@fig-thefts_map] visualizes bicycle theft counts by neighborhood across Toronto, highlighting spatial patterns in theft frequency.
 More detailed visualizations of the data are located in [@sec-data-details-visualizations]
@@
 
-The intensity of the color gradient, ranging from dark purple (low theft count) to bright yellow (high theft count), indicates the relative number of bicycle thefts in each neighborhood.
+[@fig-thefts_map] ranges from dark purple (low theft count) to bright yellow (high theft count), indicates the relative number of bicycle thefts in each neighborhood.
 The visualization reveals that bicycle thefts are highly concentrated in certain neighborhoods, particularly in the downtown core and surrounding areas, as indicated by the bright yellow and orange regions.
@@
 Our results are summarized in @tbl-modelresults.
+The intercept, representing the baseline log-odds of a theft occurring in such neighborhoods when all other predictors are held constant, is estimated at −0.386.
+This negative value suggests that, in the absence of other influencing factors, the likelihood of thefts in high-risk areas is relatively low.
+The effect of bike cost, represented as the log-transformed variable (log_bike_cost), is minimal and statistically insignificant (0.031).
+This indicates that, after accounting for premises type and time of day, the cost of the bike alone does not strongly affect the likelihood of thefts occurring in high-risk neighborhoods.
+While high-cost bikes may attract theft in general, this result implies that contextual factors, rather than cost, play a more significant role in shaping theft patterns in these areas.
+
+Premises type emerges as an important factor with notable variability in its impact.
+Theft likelihood is higher in "Commercial" (0.935) and "Educational" (0.630) premises compared to the reference category, suggesting that these locations may be more accessible or targeted by thieves.
+Conversely, thefts are less likely to occur at "House" (−1.034) and "Other" (−0.473) premises, indicating these settings may offer more security or deterrents.
+"Outside" premises (0.662) show a moderate positive association, while "Transit" (−0.575) premises display a slight negative effect, suggesting varying levels of theft risk depending on location type.
+Time of day, categorized into "Morning," "Afternoon," "Evening," and "Night," shows relatively small effects on theft likelihood.
+Compared to "Afternoon" (the reference category), "Evening" (−0.120), "Night" (−0.132), and "Morning" (−0.162) exhibit slightly reduced log-odds.
+These findings suggest minimal differences in theft risk across times of the day within high-risk neighborhoods, potentially reflecting consistent levels of vigilance or opportunity.
+
+The model’s fit, as indicated by an $R^2$ value of 0.100, shows that 10% of the variation in theft likelihood is explained by the included predictors.
+While this value suggests the presence of unaccounted-for factors, the model’s log-likelihood and information criteria (LOOIC: 1299.7, WAIC: 1299.6) indicate reasonable performance for this type of analysis.
+
+Credible intervals for the predictors, visualized in @fig-modelresults1, reveal that several premises types, such as "Commercial" and "House," show statistically significant effects, as their intervals do not cross zero.
+These results highlight the importance of contextual factors like premises type in determining theft likelihood and underscore the need for targeted strategies to mitigate risks in high-risk neighborhoods.
+
+Overall, the findings emphasize that while bike cost may not significantly influence thefts in high-risk areas, the type of premises and broader contextual factors play a pivotal role.
+These insights can inform more effective theft prevention strategies by focusing on specific high-risk premises and addressing vulnerabilities in these environments.
 
@@
 ## Model Validation
+For posterior predictive checks, @fig-post_dist demonstrates that the posterior distribution from our Bayesian logistic regression model aligns closely with the observed data on bike thefts in high-risk neighborhoods.
+This suggests that the model captures the underlying patterns of theft occurrences accurately, supporting the robustness of its predictions.
+Similarly, @fig-post_prior compares the posterior to the prior distributions, highlighting parameter changes such as those for "log_bike_cost" and "time_of_day." These shifts indicate that the observed data provided substantial information, refining our prior beliefs about these predictors.
+
+The trace plots in @fig-trace1 & @fig-trace2 show stable and horizontal chains with adequate mixing across iterations, suggesting no convergence issues in the Markov chain Monte Carlo process.
+Furthermore, the Rhat plot in @fig-rhat confirms this conclusion, as all Rhat values are close to 1 and well below the threshold of 1.05.
+This demonstrates strong convergence and reliability of the posterior estimates.
+
+Additional details and supporting figures can be found in Appendix @sec-model-details.
 

### 2024-12-03 paper/paper.qmd
discussion added

@@
 
-## First discussion point {#sec-first-point}
+## The Role of Premises in Shaping Bike Theft Risks  
+Our analysis highlights the significant influence of premises type on the likelihood of bike thefts.
+Public spaces, including "Commercial" and "Outside" premises, exhibit the highest estimated coefficients, reinforcing their status as high-risk areas.
+This finding is consistent with the assumption that public and commercial spaces, with their high foot traffic and limited security, offer increased opportunities for theft.
+Conversely, "House" premises have a negative coefficient, indicating a reduced likelihood of theft when bikes are stored at private residences.
+This reflects the higher level of control and security afforded by private properties.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+An unexpected finding pertains to "Transit" premises, which also display a negative coefficient.
+Although transit areas are often perceived as high-risk due to the volume of commuters, the results suggest that targeted measures, such as surveillance cameras and police patrols, may be deterring theft in these zones.
+However, the wider confidence intervals around this estimate indicate some uncertainty, highlighting the need for further investigation into specific security practices in transit areas.
 
-## Second discussion point
+## Timing Matters: How Time of Day Influences Bike Thefts  
+The timing of bike thefts also emerges as an important factor.
+Our model categorizes thefts into "Morning," "Afternoon," "Evening," and "Night" periods.
+Although the differences are subtler compared to premises types, theft likelihood increases slightly during "Night" hours, as reflected by its positive coefficient.
+This finding aligns with the widely held belief that lower visibility and reduced pedestrian presence at night create favorable conditions for theft.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+"Morning" and "Afternoon" periods, in contrast, exhibit lower coefficients, suggesting reduced theft activity during these times.
+This could be attributed to the higher visibility and increased presence of commuters, shoppers, and general public activity during daylight hours, which may act as a deterrent.
+These findings underscore the potential for targeted prevention measures, such as enhanced lighting and increased surveillance, particularly during evening and night hours.
 
-## Third discussion point
+## The Price of Vulnerability: Impact of Bike Cost on Theft Likelihood  
+The value of the bike, as measured by its log-transformed cost, plays a key role in predicting theft likelihood.
+The positive coefficient indicates that more expensive bikes are more likely to be targeted.
+This aligns with expectations, as high-value bikes offer greater potential resale value for thieves.
+However, the relatively small effect size suggests that factors like ease of theft and accessibility may often outweigh bike cost in determining theft likelihood.
 
-## Weaknesses and next steps
+This result underscores the need for increased awareness among bike owners, particularly those with high-value bikes, regarding secure storage practices.
+It also highlights the potential benefits of policy interventions, such as secure parking facilities in theft-prone areas, to mitigate risks associated with high-cost bicycles.
 
-Weaknesses and next steps should also be included.
+## Limitations and Challenges in Understanding Theft Patterns
+
+### Addressing the Exclusion of Location Type as a Predictor  
+Due to multicollinearity issues, location type was excluded from the final model.
+This limits the granularity of our findings, particularly in understanding theft risks associated with specific public areas, such as parks or transit stations.
+While premises type captures broader contextual differences, a more refined inclusion of location-based variables could provide a deeper understanding of the spatial dynamics of bike theft.
+
+### Balancing Generalizability Across Neighborhood Contexts  
+The model focuses on thefts occurring in high-risk neighborhoods, which may limit its applicability to areas with lower theft incidence.
+While this approach allows for targeted insights into urban centers and densely populated regions, expanding the dataset to include a more diverse range of neighborhoods could enhance generalizability and provide a fuller picture of theft dynamics across different settings.
+
+## Enhancing Prevention: Directions for Future Research and Policy  
+Future research should aim to incorporate additional contextual variables, such as weather conditions, population density, and local crime rates, to improve predictive accuracy.
+Enhanced geospatial analyses, including hot spot mapping within neighborhoods, could further refine our understanding of theft patterns.
+Moreover, investigating the effectiveness of interventions such as secure parking zones, surveillance technologies, and community engagement programs could provide actionable insights to reduce bike theft.
+These findings could inform policies and practices aimed at improving bike security and safety in urban areas.
 

### 2024-12-03 paper/paper.qmd
Survey design written

@@
 library(sf)
+library(kableExtra)
 
@@
 This format allowed for fast read-write operations and seamless integration with analysis tools.
-The cleaning and preparation steps ensured that the dataset was robust and aligned with the study’s goal of identifying theft patterns and predictors in high-risk neighborhoods.Packages used in this paper are 'tidyverse' [@tidyverse], 'here' [@here],'arrow' [@arrow], 'lubridate'[@lubridate], 'testthat' [@testthat], 'rstanarm' [@rstanarm],'sf'[@sf], 'knitr' [@knitr], and 'forcats' [@forcats].
+The cleaning and preparation steps ensured that the dataset was robust and aligned with the study’s goal of identifying theft patterns and predictors in high-risk neighborhoods.Packages used in this paper are 'tidyverse' [@tidyverse], 'here' [@here],'arrow' [@arrow], 'lubridate'[@lubridate], 'testthat' [@testthat], 'rstanarm' [@rstanarm], 'sf'[@sf], 'knitr' [@knitr], 'forcats' [@forcats] and 'kableExtra' [@kableExtra].
 
@@
 #| warning: false
+
 #### Display the first few rows ####
@@
   slice_head(n = 5) %>% # Display the first 5 rows
-  select(BIKE_COST, OCC_DATE, LOCATION_TYPE, PREMISES_TYPE) %>%
+  select(
+    `Occ Date` = OCC_DATE,
+    `Location` = LOCATION_TYPE,
+    `Premises` = PREMISES_TYPE,
+    `Bike Cost` = BIKE_COST
+  ) %>%
   kable(
     format = "markdown",
+    align = "c" # Center-align columns
   )
+
 ```
@@
 
-The bar chart shows that high-risk neighborhoods account for a significant proportion of thefts, despite representing only 10 neighborhoods out of the total.
-This concentration highlights the importance of focusing on these areas for targeted interventions.
+@fig-riskdistribution illustrates the distribution of bike thefts across different premises types within high-risk neighborhoods.
+Despite encompassing only 10 neighborhoods, these areas contribute significantly to the overall theft count, underscoring their vulnerability.
+The figure also highlights variability in premises types, with public and commercial spaces such as "Outside" and "Commercial" showing the highest theft counts, likely due to high foot traffic and lower security measures.
+Conversely, private premises like "House" exhibit fewer incidents, indicating better security controls.
+These findings emphasize the necessity of tailored intervention strategies, such as enhanced surveillance and secure bike storage options, to mitigate theft risks in these high-priority areas.
 
@@
 ## Model set-up
-The model is specified as follows: 
-\begin{align}
+
+The model is specified as follows: \begin{align}
 y_i \mid \pi_i &\sim \text{Bern}(\pi_i) \\
@@
 ## Model justification
+
 The Bayesian logistic regression was chosen for its suitability for binary outcome variables and its ability to incorporate prior information.
@@
 ## Model Validation
+
 For posterior predictive checks, @fig-post_dist demonstrates that the posterior distribution from our Bayesian logistic regression model aligns closely with the observed data on bike thefts in high-risk neighborhoods.
@@
 ## The Role of Premises in Shaping Bike Theft Risks
+
 Our analysis highlights the significant influence of premises type on the likelihood of bike thefts.
@@
 ## Timing Matters: How Time of Day Influences Bike Thefts
+
 The timing of bike thefts also emerges as an important factor.
@@
 ## The Price of Vulnerability: Impact of Bike Cost on Theft Likelihood
+
 The value of the bike, as measured by its log-transformed cost, plays a key role in predicting theft likelihood.
@@
 ### Addressing the Exclusion of Location Type as a Predictor
+
 Due to multicollinearity issues, location type was excluded from the final model.
@@
 ### Balancing Generalizability Across Neighborhood Contexts
+
 The model focuses on thefts occurring in high-risk neighborhoods, which may limit its applicability to areas with lower theft incidence.
@@
 ## Enhancing Prevention: Directions for Future Research and Policy
+
 Future research should aim to incorporate additional contextual variables, such as weather conditions, population density, and local crime rates, to improve predictive accuracy.
@@
 # Appendix {.unnumbered}
+
+# Idealized Survey Methodology
+
+## Survey Objectives
+
+The primary objective of this survey is to understand the factors influencing bike theft patterns in Toronto, with a focus on high-risk neighborhoods.
+The survey aims to capture data that cannot be directly observed through traditional theft reports, such as:
+
+-   Causal Relationships: Identifying how premises type, time of day, and bike value influence theft likelihood.
+-   Behavioral Insights: Exploring bike owners’ security habits and their perception of theft risks.
+-   Perception and Awareness: Assessing public understanding of theft risks and secure parking availability.
+-   Counterfactual Scenarios: Investigating how potential changes in infrastructure or security options might alter individual behavior.
+
+This survey is designed to directly inform strategies for targeted interventions and policy recommendations by capturing both perceptions and actual behaviors.
+By comparing responses from high-risk and low-risk neighborhoods, we aim to uncover nuanced insights into theft patterns and prevention opportunities.
+
+## Implementation Strategy
+
+### Sample Size and Target Audience
+
+The survey will focus on three key groups of participants:
+
+-   Bike Theft Victims: Individuals who have experienced theft within the last five years.
+-   Current Cyclists: Bike owners who actively use their bikes for commuting or recreation.
+-   Residents Near High- and Low-Risk Areas: To compare theft patterns across neighborhoods.
+
+The target sample size is set at **3,000 respondents**, divided as follows:
+
+-   1,000 participants from **high-risk neighborhoods**.
+-   1,000 participants from **low-risk neighborhoods**.
+-   1,000 individuals with **no prior experience of bike theft** to gather perceptions of theft risk.
+
+### Recruitment Channels
+
+To ensure diverse participation, we employ the following recruitment strategies:
+
+1.  **Digital Targeting**: Social media advertisements on platforms like Instagram, Facebook, and X (Twitter), geo-targeted to areas with reported theft incidents and high bike usage.
+2.  **Community Outreach**: Partnerships with local cycling groups, bike repair shops, and advocacy organizations to distribute the survey link.
+3.  **Email Distribution**: Outreach to residents through neighborhood associations and cycling networks.
+4.  **Incentives**: Participants receive a \$20 gift card for completing the survey.
+
+### Budget Allocation
+
+The survey has a total budget of **\$100,000**, allocated as follows:
+
+-   **\$20,000**: Digital advertising and outreach campaigns.
+-   **\$15,000**: Partnerships with cycling organizations for outreach.
+-   **\$65,000**: Participant incentives, ensuring a high response rate.
+
+## Survey Design Elements
+
+### Opening Message
+
+Participants will be welcomed with the following message:
+
+"Thank you for participating in our survey on bike thefts in Toronto.
+We aim to understand theft patterns and develop strategies to improve security for cyclists.
+Your input is invaluable and will help inform policies to reduce bike thefts.
+This survey will take approximately 5 minutes to complete, and your responses will remain anonymous.
+As a thank-you, you will receive a \$20 gift card."
+
+### Key Questions
+
+The survey includes structured questions targeting bike theft experiences and perceptions:
+
+1.  Have you experienced bike theft in the last five years?
+
+-   Yes/No
+
+2.  If yes, where did it occur?
+
+-   Home
+-   Commercial Area
+-   Transit Station
+-   Public Park
+-   Other
+
+3.  Did you report the theft to authorities?
+
+-   Yes/No/Other
+
+4.  Approximate value of the stolen bike?
+
+-   Open-ended
+
+5.  Select the word "Blue".
+
+-   Red
+-   Blue
+-   Orange
+-   Green
+
+6.  How do you secure your bike? Check all that apply:
+
+-   U-lock
+-   Chain lock
+-   Secure parking
+-   Other
+
+7.  Where do you usually park your bike?
+
+-   Street
+-   Garage
+-   Apartment Building
+-   Transit Station
+-   Other
+
+8.  On a scale of 1–5, how confident are you in your bike’s security measures?
+
+-   1 = Not confident, 5 = Very confident
+
+9.  Which time of day do you perceive as most vulnerable for bike thefts?
+
+-   Morning
+-   Afternoon
+-   Evening
+-   Night
+
+10. Which neighborhoods do you consider high-risk for bike thefts?
+
+-   Open-ended
+
+11. Are you aware of secure bike parking facilities in your area?
+
+-   Yes/No
+
+### Closing Message
+
+Participants receive a confirmation message: "Thank you for completing the survey!
+Your input will help us understand and reduce bike thefts in Toronto.
+Your \$20 gift card will be emailed within five business days.
+
+## Bias Mitigation Strategies
+
+To ensure accurate and reliable data, the survey implements the following bias mitigation strategies:
+
+1.  **Anonymity and Neutral Question Design**: Responses will remain confidential, and questions are designed to avoid leading participants toward specific answers.
+2.  **Survey Length and Engagement**: The survey is limited to 5 minutes, with a mix of question types to maintain participant interest.
+3.  **Monitoring Selection Bias**: Detailed non-response data will be collected, and statistical corrections will be applied to account for any potential bias.
+
+## Tradeoffs and Limitations
+Despite careful design, the survey faces several limitations:
+
+-   **Selection Bias**: Participants may disproportionately include individuals who have experienced theft or are particularly concerned about security.
+-   **Hypothetical Scenarios**: Responses to hypothetical questions may not fully align with real-world behaviors.
+-   **Generalizability**: Findings may reflect Toronto-specific dynamics and may not be directly applicable to other urban areas.
+
+Despite these limitations, this survey is a critical tool for exploring theft dynamics and informing effective prevention strategies.
+The insights gained will complement other data sources and contribute to a deeper understanding of bike theft patterns in Toronto.
 
@@
 ## Data visualization {#sec-data-details-visualizations}
+
 @fig-highrisk_summary is the Ranked Bar Plot for Top 10 High-Risk Neighborhoods.
@@
 ## Posterior predictive check
+
 Figure @fig-post_dist shows the posterior predictive distribution, comparing the observed data (actual theft occurrences in high-risk neighborhoods) with predictions generated by the posterior distribution of our Bayesian logistic regression model.
@@
 ## Markov chain Monte Carlo Convergence Check
+
 Figures @fig-trace1 and @fig-trace2 display the trace plots for our Bayesian logistic regression model, providing a detailed assessment of the convergence of the Markov Chain Monte Carlo (MCMC) sampling process.
@@
 This result reinforces the reliability of the posterior estimates and confirms that the sampling process has stabilized, ensuring the model’s robustness for inference.
+
 ```{r}
@@
 This visualization provides a clear summary of the model's results and highlights the most impactful predictors of theft risk.
+
 ```{r}

### 2024-12-03 paper/paper.qmd
Final edits for paper

@@
 Lastly, @sec-discussion contains discussion of the analysis based on findings, the limitations of the model and the suggestion for future research.
+@sec-survey contains information about the survey as well as its design and limitations.
+@sec-data-details contains more information about the original data and its visualizations.
+@sec-model-details shows more plots and information about the model built for this research.
 
@@
 For example, information on premises type provides insights into whether thefts are more common in public or private spaces, while bike cost highlights economic factors associated with thefts.
-A detailed list of the variables, along with sample values, is provided in the appendix (@sec-data-details) (@tbl-rawdatalist) to offer additional context for understanding the dataset.
+A detailed list of the variables, along with sample values, is provided in the appendix [@sec-data-details] (@tbl-rawdatalist) to offer additional context for understanding the dataset.
 This structured data enables a robust analysis aimed at identifying patterns and predictors of bicycle theft across Toronto.
@@
 
-# Idealized Survey Methodology
+# Idealized Survey Methodology {#sec-survey}
 
@@
 This survey will take approximately 5 minutes to complete, and your responses will remain anonymous.
-As a thank-you, you will receive a \$20 gift card."
+As a thank-you, you will receive a $20 gift card though your email within 3 business days.
+"
 
@@
 
-Participants receive a confirmation message: "Thank you for completing the survey!
+"Participants receive a confirmation message: "Thank you for completing the survey!
 Your input will help us understand and reduce bike thefts in Toronto.
-Your \$20 gift card will be emailed within five business days.
+Your \$20 gift card will be emailed within three business days.
+
+For any questions about the survey or your gift card, please contact: tommy.fu@mail.utoronto.ca ”
+
+Idealized Survey can be found here: https://forms.gle/TEMMyXt31XbYMW6c7
 
@@
 
+\newpage
 # Model details {#sec-model-details}

### 2024-12-03 other/datasheet/datasheet_template.qmd
datasheet written

---
title: "Datasheet for 'Understanding Bicycle Theft in High-Risk Neighborhoods: Key Predictors for Targeted Risk Assessment'"
subtitle: "Understanding Theft Patterns and Risks"
author: 
  - Tommy Fu
thanks: "Code and data are available at: https://github.com/YichengFu/bike_thefts_analysis.git"
date: today
date-format: long
abstract: "This datasheet documents the 'Bike Thefts Data,' a dataset aimed at understanding patterns and risks of bike thefts in urban areas. It covers the motivation, composition, collection, preprocessing, uses, and limitations of the dataset, providing critical insights for researchers, policymakers, and stakeholders interested in bike security."
format: pdf
number-sections: true
bibliography: references.bib
---

**Motivation**

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was created to explore bike theft patterns, identify high-risk areas, and understand the impact of various factors such as premises type, time of day, and bike value. It aims to inform theft prevention strategies and urban planning. Futher, the Toronto Shapefile dataset was created by the University of Toronto in the Map and GIS Library [@trt_shape].
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   The dataset was compiled by Toronto Police Open Data [@toronto_bicycle_thefts] focused on urban transportation safety and bike security.
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   No direct funded creation can be found.
4.  *Any other comments?*
    -   This informative dataset enables Bayesian regression analysis and provides insights for Theft Patterns and Risk Factors in Toronto's High-Theft Areas

**Composition**

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   The dataset represents reported incidents of bike thefts in Toronto, including information about the location (e.g., neighborhoods, premises types), time (e.g., time of day, date), bike cost, and other contextual factors. Each instance corresponds to an individual theft report.
2.  *How many instances are there in total (of each type, if appropriate)?*
    -   The dataset contains approximately 36,125 unique instances of reported bike thefts.
3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   The dataset is a sample of reported bike thefts from Toronto Police records. While it captures a broad range of thefts across neighborhoods, it may not represent unreported incidents. The geographic and temporal coverage ensures relevance for urban theft analysis.
4.  *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
    -   Each instance consists of structured data about a reported bike theft. This includes:

-   Temporal data: Occurrence date, time, and report date.
-   Geographic data: Latitude, longitude, and neighborhood identifiers.
-   Incident specifics: Type of offense, location type, premises type.
-   Bike details: Make, model, type, speed, color, and cost. Outcome: Theft status (e.g., stolen, recovered).

5.  *Is there a label or target associated with each instance? If so, please provide a description.*
    -   Yes, the STATUS column acts as a label, indicating the outcome of the theft (e.g., "STOLEN," "RECOVERED").
6.  *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
    -   Yes, some columns have missing data:
- BIKE_MODEL (many missing values, possibly because this detail wasn't reported).
- BIKE_COST and BIKE_COLOUR (some missing values).
7.  *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
    -   No explicit relationships between instances exist. Each row represents an independent theft report.
8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   There are no predefined splits in the dataset. Users may define splits based on temporal or geographic data for training/testing in predictive models.
9.  *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
    -  Possible noise includes missing or incomplete information. Some entries might have errors in reporting (e.g., incorrect dates or geographic coordinates).
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
    -   The dataset is self-contained. It does not rely on external links or resources.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
    -  No personal or confidential information about individuals appears in the dataset. It contains anonymized, aggregate-level data.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
    -   The dataset itself is not offensive. However, data related to theft and crime might be sensitive for certain users.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
    -  The dataset identifies incidents by neighborhoods but does not classify by subpopulations like age or gender.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
    -   No, the data is anonymized and does not contain personally identifiable information (PII).
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
    -   No sensitive personal data is included. It focuses on theft incidents and their characteristics.
16. *Any other comments?*
    -  The dataset is well-structured and suitable for geographic and temporal analysis of bike thefts.

**Collection process**

1.  *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
    -  Data was reported by individuals to the Toronto Police and compiled into the police records. It is directly observable and validated through police procedures.
2.  *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
    -   Data was collected through police reports, likely supplemented by digital systems for data entry.
3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   The dataset represents all reported bike thefts but excludes unreported incidents, limiting it to recorded crimes.
4.  *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
    -   The Toronto Police Service staff.
5.  *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
    -   The dataset spans multiple years. The OCC_YEAR and REPORT_YEAR columns provide temporal coverage.
6.  *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   Likely not applicable, as this is an operational dataset derived from police records.
7.  *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
    -   The data is derived directly from Toronto Police Open Data records.
8.  *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
    -   As this dataset is based on police reports, individuals were inherently aware of the data collection during the reporting process.
9.  *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
    -   By reporting thefts, individuals implicitly consented to the collection and recording of this data.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
    -   Not applicable, as the dataset anonymizes the information and focuses on incidents rather than individuals.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -  There is no indication of a formal impact analysis specific to the dataset's release.
12. *Any other comments?*
    -   The dataset appears to align with standard practices for public release of anonymized crime data.

**Preprocessing/cleaning/labeling**

1.  *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
    -   No explicit preprocessing or cleaning appears to have been performed on this dataset before release. Columns like BIKE_MODEL, BIKE_COST, and BIKE_COLOUR contain missing values. Data such as dates and times appear in a structured format but may require transformation for specific analytical purposes.
2.  *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
    -   The dataset provided seems to represent the raw or minimally processed data as collected by the Toronto Police Service. No additional raw version is referenced.
3.  *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
    -   No information about specific preprocessing or software tools is included. Users may need to conduct their own cleaning and transformation.
4.  *Any other comments?*
    -   The dataset is a suitable starting point for analysis but requires user-driven preprocessing to handle missing values and potentially derive new features for advanced tasks.

**Uses**

1.  *Has the dataset been used for any tasks already? If so, please provide a description.*
    -   No explicit mention of prior usage is included. The dataset is likely used for urban planning, crime prevention, and academic research
2.  *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
    -   There is no reference to a repository of publications or systems using this dataset.
3.  *What (other) tasks could the dataset be used for?*
    -   Potential uses:
- Predictive modeling: Identifying high-risk areas for bike theft.
- Spatial analysis: Correlation of thefts with neighborhood characteristics.
- Temporal analysis: Seasonal or hourly trends in bike theft.
- Policy evaluation: Effectiveness of theft-prevention measures.

4.  *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
    -  Missing data, especially for critical fields like BIKE_COST, may affect the quality of certain analyses. Users should address these gaps through imputation or exclusion strategies.
5.  *Are there tasks for which the dataset should not be used? If so, please provide a description.*
    -   The dataset should not be used to draw conclusions about unreported thefts or directly infer causation due to the absence of comprehensive contextual data.
6.  *Any other comments?*
    -   The dataset provides a good foundation for analyzing reported bike theft trends but requires careful consideration of its limitations.

**Distribution**

1.  *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
    -   The dataset is publicly available through the Toronto Police Open Data portal.
2.  *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
    -   Distribution occurs through an online portal, potentially as CSV files. No DOI or permanent identifier is referenced.
3.  *When will the dataset be distributed?*
    -   The dataset is already available to the public.
4.  *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
    -   The dataset is likely subject to an open data license provided by the Toronto Police Service, which permits public use.
5.  *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
    -   There are no indications of external restrictions on its use.
6.  *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
    -   No, the dataset is free from such restrictions.
7.  *Any other comments?*
    -   The dataset’s accessibility supports transparency and public engagement with municipal crime data.

**Maintenance**

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   The Toronto Police Service is responsible for maintaining the dataset.
2.  *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
    -   Contact details for the Toronto Police Service’s Open Data available on their official website.
3.  *Is there an erratum? If so, please provide a link or other access point.*
    -   No errata have been referenced or linked.
4.  *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
    -  Updates are based on periodic data releases by the Toronto Police Service. The frequency and notification mechanism are unclear.
5.  *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
    -   No explicit retention policy is indicated, but anonymization ensures compliance with privacy standards.
6.  *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
    -   No information is provided, though users are encouraged to access the latest version from the Toronto Police Open Data portal.
7.  *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
    -   The dataset does not support direct contributions but allows users to conduct independent analyses.
8.  *Any other comments?*
    -   Clear documentation and periodic updates would enhance the dataset’s usability.

\newpage

# References

### 2024-12-03 paper/paper.qmd
Final Edits

@@
 toc: true
-abstract: "This study explores patterns of bicycle theft in Toronto, focusing on thefts in high-risk neighborhoods using logistic regression.
-The analysis identifies bike cost, premises type, and time of occurrence as significant predictors of thefts in high-risk neighborhoods.
-Results show that high-cost bikes are disproportionately stolen in public spaces and during evening hours, with specific neighborhoods exhibiting elevated theft rates.
-These findings highlight risk factors and provide actionable insights for targeted interventions, urban planning, and improved security measures in vulnerable areas."
+abstract: "This analysis investigates bicycle theft in Toronto, emphasizing high-risk neighborhoods.
+Using Bayesian logistic regression, the study identifies key predictors, including premises type, bike cost, and time of day.
+Results reveal that theft likelihood increases in public spaces and during evening hours, with high-cost bicycles as prime targets.
+Specific neighborhoods consistently report elevated theft rates, suggesting localized vulnerabilities.
+These findings support targeted interventions, such as improved security measures and community awareness, to mitigate theft risks and enhance urban safety."
 format: pdf

### 2024-12-03 paper/paper.qmd
Final Edits

@@
 
-This research provides actionable knowledge to enhance theft prevention in urban settings.
-By identifying the circumstances under which thefts are more likely to occur, the study offers evidence to inform targeted interventions, such as improved security measures in vulnerable areas.
-These findings contribute to creating safer urban spaces for cyclists.
+Bicycle theft is a persistent issue in urban environments, imposing economic and emotional burdens on individuals and communities.
+As cities like Toronto increasingly embrace cycling as a sustainable transportation mode, mitigating theft is essential for maintaining public trust in cycling infrastructure.
+However, existing studies often focus on general crime trends, overlooking the nuanced factors that drive bicycle theft in specific high-risk neighborhoods.
+
+This study addresses this gap by analyzing reported bicycle thefts in Toronto.
+Using statistical modeling, it identifies contextual factors such as premises type, time of occurrence, and bike cost that influence theft likelihood in high-risk neighborhoods.
+By focusing on these areas, the research highlights actionable risk factors and proposes strategies to prevent theft and enhance cyclist security.
 
@@
 
+The results of the Bayesian logistic regression model emphasize the role of contextual and temporal factors in shaping bicycle theft patterns. Key findings include:
+
+- Premises Type: Theft risk is significantly higher in public spaces such as streets and parks compared to residential premises, likely due to lower security measures.
+- Time of Day: The likelihood of theft increases during evening hours, reflecting reduced visibility and activity.
+- Bike Cost: High-cost bicycles are more frequently targeted, underscoring their appeal to thieves.
+
 Premises type emerges as an important factor with notable variability in its impact.
@@
 It also highlights the potential benefits of policy interventions, such as secure parking facilities in theft-prone areas, to mitigate risks associated with high-cost bicycles.
+
+## Implications for Urban Policy
+These findings suggest actionable strategies for reducing bicycle theft in urban areas by addressing vulnerabilities in public spaces.
+High-risk locations, such as streets, parks, and transit areas, often lack adequate security measures, making them prime targets for theft.
+Implementing targeted interventions, such as enhanced lighting, can deter criminal activity by increasing visibility and discouraging opportunistic behavior.
+Community policing, involving regular patrols and partnerships with local cycling groups, can create a sense of security and deter theft through increased accountability.
+Secure parking zones equipped with features such as CCTV surveillance, well-lit areas, and designated bike racks can provide cyclists
 
